---
title: "Quarto setup"
description: "Guidance on Installing and Configuring Quarto"
author: "RPythonStudyGroup feat. ChatGPT"
date: "2024-10-24"
date-modified: "2024-10-24"
categories:
  - Quarto
  - setup
---

:::: {.content-visible when-profile="ko"}
## 설치안내문서
Quarto 공식사이트에서 제공하는 공식문서(<https://quarto.org/docs/get-started/>)를 참고하는 것이 이상적이겠지만, 아래의 요약을 참고하여 설치 하셔도 됩니다.

## 설치 파일
공식 Python 웹사이트(<https://quarto.org/docs/get-started/>)에서 운영체제에 적합한 최신 Quarto 설치파일을 다운로드하고 설치합니다. 2024년 10월 24일 기준으로 Windows 운영 체제의 경우 `quarto-1.5.557-win.msi` 설치파일(2024년 8월 30일 release)이 최신입니다.

## 설치옵션

### 설치경로

윈도우의 경우 기본 설치경로인 `C:\Program Files\Quarto\`를 추천합니다.
사용하거나 사용자가 원하는 경로를 선택할 수 있습니다. (시스템의 환경 변수(PATH) 등록은 자동으로 이루어 집니다.)


## 설치 확인

::: {.callout-note title="Python이 올바르게 설치되었는지 확인하려면," collapse="true" appearance="minimal"}
Win + R을 눌러 실행 대화창을 열고 `cmd`를 입력하여 명령 프롬프트를 실행합니다.

```{r cmd_kr, eval=FALSE, filename="Command Prompt"}
cmd
```

명령 프롬프트에서 다음 명령을 입력하여 설치된 quarto 버전이 예상 버전과 일치하는지 확인합니다.

```{r quarto_version_kr, eval=FALSE, filename="Command Prompt"}
quarto --version
```

:::
::::

:::: {.content-visible when-profile="en"}

## Installation Guide
You can refer to the official documentation provided by Quarto at <https://quarto.org/docs/get-started/>. However, you can also follow the summary below to install Quarto.

## Installation File
Download the latest Quarto installation file suitable for your operating system from the official Python website (<https://quarto.org/docs/get-started/>) and install it. As of October 24, 2024, the `quarto-1.5.557-win.msi` installation file (released on August 30, 2024) is the latest for Windows operating systems.

## Installation Options

### Installation Path

For Windows, it is recommended to use the default installation path `C:\Program Files\Quarto\`.
You can also choose to use any other path you prefer. (The system environment variable (PATH) registration is done automatically.)

## Verification of Installation

::: {.callout-note title="To verify if Python is installed correctly," collapse="true" appearance="minimal"}
Press Win + R to open the Run dialog box and type `cmd` to run the command prompt

```{r cmd_en, eval=FALSE, filename="Command Prompt"}
cmd
```

In the command prompt, enter the following command to check if the installed quarto version matches the expected version.

```{r quarto_version_en, eval=FALSE, filename="Command Prompt"}
quarto --version
```

:::
::::